### 2024-11-12 Paper/Article.Rmd
Clean up and add figures

---
title: "Just Take the Bus: Analysis of Toronto Bike Theft from 2014-2023"
subtitle: "Examining Patterns, Locations, and Factors of Bicycle Theft in Toronto"
author:
- Rayan Awad Alim
- Samreen Gill
thanks: https://github.com/RayanAlim/Toronto_Bike_Thefts_Data_Analysis
date: "November 12, 2024"
abstract: This analysis examines Bike theft data obtained from the Tronrot Police Data Portal.
output: pdf_document
toc: true
number-sections: true
editor_options:
  markdown:
    wrap: 72
---

```{r setup}
#| include: false
#| warning: false
#| message: false

# Workspace Set up

knitr::opts_chunk$set(echo=FALSE, include=TRUE, eval=TRUE)
required_packages <- c(
  "lubridate",
  "sf",
  "tidyverse",
  "ggplot2",
  "maps",
  "dplyr",
  "mapdata",
  "ggspatial",
  "opendatatoronto",
  "here",
  "stringr",
  "RColorBrewer",
  "Scales"
) 
for (p in required_packages) { # install if missing
  if (!require(p, character.only = TRUE)) {
    install.packages(p, character.only = TRUE)
  }
}

library(lubridate)
library(sf)
library(tidyverse)
library(ggplot2)
library(maps)
library(dplyr)
library(mapdata)
library(ggspatial)
library(opendatatoronto)
library(here)
library(stringr)
library(RColorBrewer)
library(scales)
```

```{r}
#| include: false
#| warning: false
#| message: false

# Load and process data from CSV and GeoJSON files
bike_thefts <- read_csv(here("Data", "Bike_thefts_raw.csv")) |>
  mutate(
    occurence_date = as_date(
      OCC_DATE, format = "%a, %d %b %Y %H:%M:%S GMT"
    ),
    report_date = as_date(
      REPORT_DATE, format = "%a, %d %b %Y %H:%M:%S GMT"
    ),
    NEIGHBOURHOOD_Clean = str_remove(NEIGHBOURHOOD_158, "\\s*\\(\\d+\\)$")
  )

bike_thefts_sf <- read_sf(here("Data", "Bicycle_Thefts_Open_Data.geojson")) |>
  mutate(
    occurence_date = as_date(
      OCC_DATE, format = "%a, %d %b %Y %H:%M:%S GMT"
    ),
    report_date = as_date(
      REPORT_DATE, format = "%a, %d %b %Y %H:%M:%S GMT"
    ),
    NEIGHBOURHOOD_Clean = str_remove(NEIGHBOURHOOD_158, "\\s*\\(\\d+\\)$")
  )

toronto_shape <- read_sf(here("Data", "Neighbourhoods - 4326.shp")) 
```

```{r}
# glimpse(toronto_shape)
# glimpse(bike_thefts)
```

# Introduction

Many cities are promoting bicycles to reduce reliance on cars, recognizing significant benefits cycling brings to the environment, public health and economy. As cyclists continue to gain popularity in Toronto, bicycle theft has emerged as a growing concern leaving many feeling vulnerable. With the increasing number of cyclists on the road, knowing when and where bike thefts are most likely to occur is crucial for protecting valuable bicycles. This analysis aims to inform the general public and cyclists about neighborhoods with the highest rate of thefts and times when their bikes are most likely to be stolen. This would not only raise awareness but also help people make more informed decisions to secure their bikes, ultimately reducing the risk of theft and fostering a safer cycling environment in Toronto.

**Top 10 neighborhoods that contribute to the most number of bicycle thefts across Toronto ?**

We can see the worst neighborhood for bike thefts in Toronto is Yonge-Bay Corridor followed by Downtown Yonge East and St Lawrence East Bayfront, with each exceeding 1000 incidents from 2022 to 2024. In the past three years, about 5100 bikes were reported stolen from these areas. These neighborhoods contribute to approximately 17% of total bicycle thefts reported across Toronto. This high concentration reflects a combination of factors such as lack of secure bike storage, crowded public spaces making these areas prime targets for thieves.

How much do they contribute to total theft from 2022 to 2024 ? (Comparing with premise type)An in-depth analysis of premise type unfolds that bicycles last parked outside building premises, whether residential or commercial are particularly vulnerable to theft. This trend is consistently observed in top 10 neighborhoods for bike theft. These locations, especially because of high foot or bike traffic, provide easy access and quick escape routes for criminals.

```{r}
# Samreen 
# bar charts of general statistics:
# # of bikes stole in 2021
# # of bikes stole in 2022
# # of bikes stole in 2023
# Total Loss of x $ M 
```

```{r}
# colnames(toronto_shape)
# colnames(thefts_by_neighborhood)
```

```{r}
thefts_by_neighborhood <- bike_thefts %>%
  count(NEIGHBOURHOOD_Clean) %>%
  rename(thefts_count = n)

toronto_shape <- toronto_shape %>%
  rename(NEIGHBOURHOOD_Clean = AREA_NA7)

toronto_thefts_map <- toronto_shape %>%
  left_join(thefts_by_neighborhood, by = "NEIGHBOURHOOD_Clean")

ggplot(toronto_thefts_map) +
  geom_sf(aes(fill = thefts_count), color = "grey", size = 0.1) +
  scale_fill_viridis_c(option = "plasma", name = "Number of Thefts") +
  labs(title = "Number of Bike Thefts per Neighborhood in Toronto") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))
```
premise type:

```{r}
bike_thefts %>%
  count(PREMISES_TYPE) %>%
  ggplot(aes(x = "", y = n, fill = PREMISES_TYPE)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  labs(title = "Bike Theft by Location Type",
       fill = "Location Type") +
  theme_void()
```

## When is your bike most likely to be stolen(day of the week, month) ?

Based on the figures below, it is pretty evident the highest frequency of bicycle thefts occurs during July and August. This aligns with the warmer summer months when more people tend to ride bicycles, increasing both bicycles in public spaces and likelihood of theft. Thefts begin to build up during spring, peaking in summer, followed by a steady decline through fall and winter. The data suggests thefts are more common during weekdays, potentially because most people use bicycles to commute to work or school. Bicycles parked during weekdays are often unattended for several hours, providing thieves with ample time to steal them.

Additionally, the majority of bicycles get stolen during the night than in morning, afternoon or evening.

```{r}
# Bike Thefts by Month
bike_thefts <- bike_thefts %>%
  mutate(OCC_MONTH = factor(OCC_MONTH, levels = month.name))

bike_thefts %>%
  count(OCC_MONTH) %>%
  ggplot(aes(x = OCC_MONTH, y = n, fill = OCC_MONTH)) +
  geom_bar(stat = "identity") +
  labs(title = "Bike Thefts by Month of the Year",
       x = "Month", y = "Number of Thefts", fill = "Months") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

```{r}
bike_thefts <- bike_thefts %>%
  mutate(OCC_DOW = factor(OCC_DOW, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")))

bike_thefts %>%
  count(OCC_DOW) %>%
  ggplot(aes(x = OCC_DOW, y = n, fill = OCC_DOW)) +
  geom_bar(stat = "identity") +
  labs(title = "Bike Theft Patterns by Week Day Slight Uptick on Fridays",
       x = "Day of the Week", 
       y = "Number of Thefts", 
       fill = "Day of Week",
       caption = "Bike thefts in Toronto show a consistent pattern across weekdays, with a small increase on Fridays.") +  
  theme_minimal() +
  scale_fill_brewer(palette = "Set3")

```

```{r}
bike_thefts <- bike_thefts %>%
  mutate(OCC_DOW = factor(OCC_DOW, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")))

ggplot(bike_thefts, aes(x = OCC_DOW, y = OCC_HOUR)) +
  geom_bin2d(bins = 24) +
  scale_y_reverse() + 
  scale_fill_gradient(low = "lightblue", high = "red", name = "Number of Thefts") +  
  labs(
    title = "Bike Theft Patterns by Day and Hour",
    x = "Day of the Week",
    y = "Hour of Day",
    caption = "TBD",
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

### Neighbourhood analysis and Maps

```{r}
bike_thefts %>%
  count(NEIGHBOURHOOD_Clean) %>%
  slice_max(n, n = 10) %>%
  ggplot(aes(x = reorder(NEIGHBOURHOOD_Clean, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = n), hjust = 1.2, color = "white") +
  coord_flip() +
  labs(title = "Downtown Danger: Top 10 Neighborhoods for Bike Thefts",
       x = "Neighborhood", y = "Number of Thefts") +
  theme_minimal()
```

```{r}
bike_thefts %>%
  count(NEIGHBOURHOOD_Clean) %>%
  arrange(desc(n)) %>%
  slice_max(n, n = 10) %>%
  ggplot(aes(x = "", y = n, fill = reorder(NEIGHBOURHOOD_Clean, -n))) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  geom_text(aes(label = paste0("(", round(n / sum(n) * 100, 1), "%)")),
            position = position_stack(vjust = 0.5),
            color = "white", size = 4) +
  labs(title = "Top 10 Neighborhoods for Bike Thefts",
       fill = "Neighborhood") +
  scale_fill_brewer(palette = "Paired") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

```

### Bike Theft by Type and Cost

```{r}
total_cost_by_neighborhood <- bike_thefts %>%
  group_by(NEIGHBOURHOOD_Clean) %>%
  summarize(total_cost = sum(BIKE_COST, na.rm = TRUE))

toronto_cost_map <- toronto_shape %>%
  left_join(total_cost_by_neighborhood, by = c("NEIGHBOURHOOD_Clean" = "NEIGHBOURHOOD_Clean"))

ggplot(toronto_cost_map) +
  geom_sf(aes(fill = total_cost), color = "grey", size = 0.1) +
scale_fill_viridis_c(option = "inferno", name = "Total Theft Cost ($)",
                       breaks = seq(0, max(toronto_cost_map$total_cost, na.rm = TRUE), length.out = 10),
                       labels = scales::comma_format()) + 
  labs(title = "Total Cost of Bike Thefts per Neighborhood in Toronto") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))
```

Here we look at the frequency of thefts by bike type:\n
```{r}
total_cost_by_neighborhood %>%
  arrange(desc(total_cost)) %>%
  slice_max(total_cost, n = 10) %>%  
  ggplot(aes(x = reorder(NEIGHBOURHOOD_Clean, total_cost), y = total_cost)) +
  geom_bar(stat = "identity", fill = "tan4") +
  geom_text(aes(label = scales::comma(total_cost)), hjust = 1.2, color = "white") +
  coord_flip() +
  labs(title = "Top 10 Neighborhoods by Total Theft Cost",
       x = "Neighborhood", y = "Total Theft Cost ($)") +
  theme_minimal()
```

```{r}
bike_type_names <- c(
  "BM" = "BMX", 
  "EL" = "Electric Bike", 
  "FO" = "Folding Bike", 
  "MT" = "Mountain Bike", 
  "OT" = "Other Bike", 
  "RC" = "Road Bike", 
  "RE" = "Hybrid Bike",
  "RG" = "??", 
  "SC" = "??", 
  "TA" = "??", 
  "TO" = "??", 
  "TR" = "??", 
  "UN" = "Unspecified"
)

custom_colors <- c("royalblue", "maroon3", "springgreen4", "salmon",
                   "purple4", "wheat2", "maroon", "seagreen1", "khaki4",
                   "lightseagreen", "#9b59b6", "chocolate1", "red3")

ggplot(bike_thefts, aes(x = BIKE_TYPE, fill = BIKE_TYPE)) +
  geom_bar() +
  labs(
    title = "Number of Bike Thefts by Bike Type",
    x = "Bike Type",
    y = "Count",
    fill = "Bike Type"  
  ) +
  scale_fill_manual(values = custom_colors, labels = bike_type_names) +  
  theme_minimal() +
  theme(
    legend.position = "right",  
    axis.text.x = element_text(angle = 45, hjust = 1)  
  )
```

### Reporting Delays

Analyzing the time delay between theft occurrence and reporting date
(Not working, got to tweajk it some more)

```{r}
# Pie Chart of recovery vs stolen

recovery_status <- bike_thefts %>%
  count(STATUS) %>%
  mutate(percentage = n / sum(n) * 100)

ggplot(recovery_status, aes(x = "", y = n, fill = STATUS)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), 
            position = position_stack(vjust = 0.5),
            color = "white") +
  labs(title = "Bike Recovery Status: Recovered vs Stolen",
       fill = "Status") +
  theme_void() +
  scale_fill_brewer(palette = "Dark2")

```
```{r}
# MAP OF RECOVERY RATE 
recovery_by_neighborhood <- bike_thefts %>%
  group_by(NEIGHBOURHOOD_Clean) %>%
  summarize(
    total_thefts = n(),
    recovered = sum(STATUS == "RECOVERED", na.rm = TRUE),
    recovery_rate = recovered / total_thefts
  )

if (!"sf" %in% class(recovery_by_neighborhood)) {
  toronto_recovery_map <- toronto_shape %>%
    left_join(recovery_by_neighborhood, by = "NEIGHBOURHOOD_Clean")
} else {
  toronto_recovery_map <- st_join(toronto_shape, recovery_by_neighborhood, left = TRUE)
}

ggplot(toronto_recovery_map) +
  geom_sf(aes(fill = recovery_rate), color = "grey") +
  scale_fill_viridis_c(option = "plasma", name = "Recovery Rate") +
  labs(title = "Bike Theft Recovery Rate by Neighborhood in Toronto") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))
```

### Theft Recovery Analysis by Neighborhood

which neighborhoods have higher recovery rates and analyzing the
recovered vs unrecovered bikes by neighborhood

```{r}
# calculating recovery rate by neighborhood
recovery_by_neighborhood <- bike_thefts %>%
  group_by(NEIGHBOURHOOD_Clean) %>%
  summarize(
    total_thefts = n(),
    recovered = sum(STATUS == "RECOVERED"),
    recovery_rate = recovered / total_thefts
  ) %>%
  slice_max(recovery_rate, n = 10)

ggplot(recovery_by_neighborhood, aes(x = reorder(NEIGHBOURHOOD_Clean, recovery_rate), y = recovery_rate)) +
  geom_bar(stat = "identity", fill = "forestgreen") +
  coord_flip() +
  labs(title = "Top 10 Neighborhoods by Recovery Rate", x = "Neighborhood", y = "Recovery Rate") +
  theme_minimal()
```