### 2024-11-13 Paper/Article.Rmd
Update paper

---
title: "Just Take the Bus: Analysis of Toronto Bike Theft from 2014-2023"
subtitle: "Examining Patterns, Locations, and Factors of Bicycle Theft in Toronto"
author:
- Rayan Awad Alim & Samreen Kaur
thanks: https://github.com/RayanAlim/Toronto_Bike_Thefts_Data_Analysis
# date: "November 12, 2024"
# abstract: This analysis examines Bike theft data obtained from the Tronrot Police Data Portal.
output: pdf_document
editor_options:
  markdown:
    wrap: 72
---

```{r setup}
#| include: false
#| warning: false
#| message: false
rm(list=ls())
options(repos = c(CRAN = "https://cran.rstudio.com/"))

knitr::opts_chunk$set(echo=FALSE, include=TRUE, eval=TRUE)
required_packages <- c(
  "lubridate",
  "sf",
  "tidyverse",
  "ggplot2",
  "maps",
  "dplyr",
  "mapdata",
  "ggspatial",
  "opendatatoronto",
  "here",
  "stringr",
  "RColorBrewer",
  "Scales",
  "pak",
  "patchwork",
  "gridExtra",
  "ggrepel"
) 
for (p in required_packages) { # install if missing
  if (!require(p, character.only = TRUE)) {
    install.packages(p, character.only = TRUE)
  }
}

library(lubridate)
library(sf)
library(tidyverse)
library(ggplot2)
library(maps)
library(dplyr)
library(mapdata)
library(ggspatial)
library(opendatatoronto)
library(here)
library(stringr)
library(RColorBrewer)
library(scales)
library(pak)
library(patchwork)
library(gridExtra)
library(ggrepel)
```

```{r}
#| include: false
#| warning: false
#| message: false

# Load and process data from CSV and GeoJSON files
bike_thefts <- read_csv(here("Data", "Bike_thefts_raw.csv")) |>
  mutate(
    occurence_date = as_date(
      OCC_DATE, format = "%a, %d %b %Y %H:%M:%S GMT"
    ),
    report_date = as_date(
      REPORT_DATE, format = "%a, %d %b %Y %H:%M:%S GMT"
    ),
    NEIGHBOURHOOD_Clean = str_remove(NEIGHBOURHOOD_158, "\\s*\\(\\d+\\)$")
  )

bike_thefts_sf <- read_sf(here("Data", "Bicycle_Thefts_Open_Data.geojson")) |>
  mutate(
    occurence_date = as_date(
      OCC_DATE, format = "%a, %d %b %Y %H:%M:%S GMT"
    ),
    report_date = as_date(
      REPORT_DATE, format = "%a, %d %b %Y %H:%M:%S GMT"
    ),
    NEIGHBOURHOOD_Clean = str_remove(NEIGHBOURHOOD_158, "\\s*\\(\\d+\\)$")
  )

toronto_shape <- read_sf(here("Data", "Neighbourhoods - 4326.shp")) 
```

If you live in Toronto, brace yourself: if your bike gets stolen, the chances of getting it back are slim. In fact, the odds are roughly 1%- about as likely as your chances at being struck by lightning twice in your Life. But don’t give up hope just yet! In this data-driven analysis, we break down the patterns and factors behind bike theft in Toronto, so cyclists can arm themselves with knowledge, and maybe a stronger lock.

This analysis aims to inform the general public and cyclists about Toronto’s bike theft hotspots, highlighting neighborhoods with the highest rate of thefts and times when bikes are most at risk and recovery rate of different neighbourhoods. It would not only raise awareness but also help general public and Toronto police take proactive measures to secure bikes, ultimately reducing the risk of theft and fostering a safer cycling environment in Toronto.

```{r}
#| fig.width: 6
#| fig.height: 7
#| fig.align: center

# Pie Chart of recovery vs stolen

recovery_status <- bike_thefts %>%
  count(STATUS) %>%
  mutate(percentage = n / sum(n) * 100,
         label = paste0(STATUS, ": ", round(percentage, 1), "%"))

ggplot(recovery_status, aes(x = "", y = n, fill = STATUS)) +
  geom_bar(stat = "identity", width = 0.6) +
  coord_polar("y", start = 0) +
  geom_label_repel(aes(label = label, fill = STATUS),            
                   color = "white",                              
                   size = 4,                                     
                   nudge_x = 0.8,                                
                   show.legend = FALSE,                          
                   segment.color = "grey20",                    
                   segment.size = 0.6) + 
  labs(title = "Bike Recovery Status: Recovered vs. Stolen") +
  theme(legend.position = "none") +
  theme_void() +
  scale_fill_brewer(palette = "Dark2") 
```

### Bicycle Theft At a Steady Increase Post-pandemic

Analyzing trend in bicycle theft over past 5 years, notice that there is
approximately 15% drop in bike thefts from 2019 to 2023. This suggests
bike theft was a significant issue prior to COVID-19 pandemic. During
2020 and 2021, decline in thefts could be associated with lock down and
reduced outdoor activities since fewer people were commuting and leaving
their bikes in public spaces. The gradual stabilization in subsequent
years(2022 and 2023) reflects either became more cautious about security
of their bikes or thieves shifted their focus to other crimes, making
theft less prevalent in these neighborhoods.

```{r}
# total thefts by year in top 10 neighborhoods

bike_thefts %>%
  filter(OCC_YEAR %in% c(2019, 2020, 2021, 2022, 2023)) %>%
  group_by(OCC_YEAR, NEIGHBOURHOOD_Clean) %>%
  count() %>%
  ungroup() %>%
  group_by(OCC_YEAR) %>%
  top_n(10, n) %>%
  summarise(total_thefts = sum(n)) %>%
  ggplot(aes(x = factor(OCC_YEAR), y = total_thefts, group = 1)) + 
  geom_line(color = "#458CFF") +
  geom_point(size = 3) +
  labs(title = "Yearly Breakdown of Bicycle Theft Trends in Toronto",
       x = "Year", y = "Total Thefts") +
  theme_minimal() +
  theme(legend.position = "none")
```

### Your bike is most likely to be stolen in Downtown

The worst neighborhood for bike thefts in Toronto is Yonge-Bay Corridor
followed by Downtown Yonge East and St Lawrence East Bayfront, with each
exceeding 1000 incidents over last 5 years. In the past three years,
about 5100 bikes were reported stolen from these areas. These
neighborhoods contribute to approximately 17% of total bicycle thefts
reported across Toronto. This high concentration reflects a combination
of factors such as lack of secure bike storage, crowded public spaces
making these areas prime targets for thieves.

Among these incidents, bicycles are predominantly stolen outside
building premises, whether residential or commercial. This trend is
consistently observed in top 10 neighborhoods for bike theft. These
locations, especially because of high foot or bike traffic, provide easy
access and quick escape routes for criminals. Moreover outdoor spaces
often lack surveillance cameras, making it easier for thieves to act
without being noticed. Similarly, bike storage areas in apartments, such
as shared parking garages or storage rooms, offer limited protection due
to communal access resulting in high volume of thefts. The visualization
below shows that bikes are safe when bikers are at school or in transit.
Bikers association can use these findings to educate bikers about safe
storage practices especially in outdoor spaces and apartments. City
planners should in coordination with apartment managers to install
secure bike storage facilities and ensure public bike racks are located
in well-monitored, well-lit locations.

```{r}
plot_1<- thefts_by_neighborhood <- bike_thefts %>%
  count(NEIGHBOURHOOD_Clean) %>%
  rename(thefts_count = n)

toronto_shape <- toronto_shape %>%
  rename(NEIGHBOURHOOD_Clean = AREA_NA7)

toronto_thefts_map <- toronto_shape %>%
  left_join(thefts_by_neighborhood, by = "NEIGHBOURHOOD_Clean")

 plot_1<-ggplot(toronto_thefts_map) +
  geom_sf(aes(fill = thefts_count), color = "grey", size = 0.1) +
  scale_fill_viridis_c(option = "plasma", name = "No. of Thefts") +
  labs(caption = "Thefts per Neighborhood in Toronto") +
  theme_void() +
  theme(plot.caption = element_text(hjust = 0.5))

plot_2<- bike_thefts %>%
  count(PREMISES_TYPE) %>%
   ggplot(aes(x = "", y = n, fill = PREMISES_TYPE)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  labs(caption = "Thefts by Location Type",
       fill = "Location Type") +
  theme_void()

grid.arrange(plot_1, plot_2, ncol=2, widths = c(2, 2))

```

### Seasonal and weekly patterns of bicycle theft

The highest frequency of bicycle thefts occurs during July and August.
This aligns with the warmer summer months when more people tend to ride
bicycles, increasing both bicycles in public spaces and their likelihood
of theft. Thefts begin to build up during spring, peaking in summer,
followed by a steady decline through fall and winter. The data suggests
thefts are more common during weekdays, potentially because most people
use bicycles to commute to work or school. Bicycles parked during
weekdays are often unattended for several hours, providing thieves with
ample time to steal them. On weekends, cyclists may spend more time with
their bikes using them for recreational purposes, which means bicycles
are less likely to be left unattended in public spaces leading to slight
decrease in thefts during weekends. This shows Toronto police should
allocate more resources in downtown particularly during weekdays to
reduce the frequency of theft in high-risk areas.

```{r}
#| fig.width: 10
#| fig.height: 7
#| fig.align: center
#| 
# Bike Thefts by Month
bike_thefts <- bike_thefts %>%
  mutate(OCC_MONTH = factor(OCC_MONTH, levels = month.name))

plot_by_month <- bike_thefts %>%
  count(OCC_MONTH) %>%
  ggplot(aes(x = OCC_MONTH, y = n, fill = OCC_MONTH)) +
  geom_bar(stat = "identity") +
  labs(title = "Seasonal Thefts, Nothing Like A Sunny \n Day To Snatch a Bike",
       caption = "Bike Thefts by Month of the Year",
       x = "Month", y = "Number of Thefts") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill = "none") +
  theme(plot.caption = element_text(hjust = 0.5))

bike_thefts <- bike_thefts %>%
  mutate(OCC_DOW = factor(OCC_DOW, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")))

plot_by_day <- bike_thefts %>%
  count(OCC_DOW) %>%
  ggplot(aes(x = OCC_DOW, y = n, fill = OCC_DOW)) +
  geom_bar(stat = "identity") +
  labs(title = "Bike Theft Patterns by Week Day Slight\n Uptick on Fridays",
       x = "Day of the Week", 
       y = "Number of Thefts", 
       fill = "Day of Week",
       caption = "A consistent pattern across weekdays, with a small increase on Fridays.") +  
  theme_minimal() +
  scale_fill_brewer(palette = "Set3") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill = "none")

grid.arrange(plot_by_month, plot_by_day, ncol=2, widths = c(3, 3))

```

### The Hidden Danger: Nighttime bicycle thefts in Downtown

Bike thefts are more prevalent during night, primarily due to the cover
of darkness thieves couldn't be easily detected. In downtown, during
weekdays these neighborhoods tend to be less crowded and there are fewer
people to notice any suspicious activity. Although people leave there
bikes locked overnight, thieves tend to break locks or steal parts of
bikes to resell them. This calls for improved lighting and overnight
surveillance to deter thieves during these vulnerable hours.

```{r}
#| fig.height: 3
#| fig.align: center

bike_thefts <- bike_thefts %>%
  mutate(OCC_DOW = factor(OCC_DOW, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")))

ggplot(bike_thefts, aes(x = OCC_DOW, y = OCC_HOUR)) +
  geom_bin2d(bins = 24) +
  scale_y_reverse() + 
  scale_fill_gradient(low = "lightblue", high = "red", name = "Number of Thefts") +  
  labs(
    title = "Bike Theft Patterns by Day and Hour",
    x = "Day of the Week",
    y = "Hour of Day"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

### Toronto's Bike theft economy: Hidden costs for all

Bicycle theft in Toronto has a broader economic impact beyond the direct
cost to individual victims. It not only affects personal finances of
bikers but also the local economy and public perception of safety in the
city. Bike thefts deter people from using bikes as primary mode of
transport affecting businesses like bike shops, repair services that
solely rely on cyclists as customers. Moreover, with thousands of thefts
being reported annually requires police resources for investigation and
record-keeping. This strain on law enforcement can becomes costly,
ultimately impacting taxpayers when more resources are allocated to
reduce thefts.

```{r}
#| fig.height: 3
#| fig.align: center

total_cost_by_neighborhood <- bike_thefts %>%
  group_by(NEIGHBOURHOOD_Clean) %>%
  summarize(total_cost = sum(BIKE_COST, na.rm = TRUE))

toronto_cost_map <- toronto_shape %>%
  left_join(total_cost_by_neighborhood, by = c("NEIGHBOURHOOD_Clean" = "NEIGHBOURHOOD_Clean"))

ggplot(toronto_cost_map) +
  geom_sf(aes(fill = total_cost), color = "grey", size = 0.1) +
scale_fill_viridis_c(option = "inferno", name = "Total Theft Cost ($)",
                       breaks = seq(0, max(toronto_cost_map$total_cost, na.rm = TRUE), length.out = 10),
                       labels = scales::comma_format()) + 
  labs(title = "Total Cost of Bike Thefts per Neighborhood in Toronto") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))
```

### Mountain and Regular bikes are the most stolen bikes

Mountain and regular bikes are most commonly stolen, with each group
accounting for 30% of overall theft. The high theft rates for these
types may be attributed to their popularity and general accessibility in
Toronto, as they both are commonly used for commuting and recreational
purposes in Toronto. Theft of electric bikes are relatively lower in
comparison. This may be because of increased weight of bikes, or added
security features often built into them. Scooters and other unique types
of bicycles like tandem, tricycles, and unicycles show very low theft
rates, likely because they are less commonly used. Bike association
should encourage bike owners, especially those with high-risk models
like Mountain and Regular bikes, to install GPS trackers. These devices
can be discreetly attached to bikes and synced with smartphone apps,
enabling real-time tracking if the bike is stolen.

```{r}
#| fig.height: 3
#| fig.align: center

bike_type_names <- c(
  "BM" = "BMX", 
  "EL" = "Electric Bike", 
  "FO" = "Folding Bike", 
  "MT" = "Mountain Bike", 
  "OT" = "Other Bike", 
  "RC" = "Road Bike", 
  "RE" = "Hybrid Bike",
  "RG" = "Regular Bike", 
  "SC" = "Scooter", 
  "TA" = "Tandem", 
  "TO" = "Touring", 
  "TR" = "Tricycle", 
  "UN" = "Unicycle"
)

custom_colors <- c("royalblue", "maroon3", "springgreen4", "salmon",
                   "purple4", "wheat2", "maroon", "seagreen1", "khaki4",
                   "lightseagreen", "#9b59b6", "chocolate1", "red3")

ggplot(bike_thefts, aes(x = BIKE_TYPE, fill = BIKE_TYPE)) +
  geom_bar() +
  labs(
    title = "Number of Bike Thefts by Bike Type",
    x = "Bike Type",
    y = "Count",
    fill = "Bike Type"  
  ) +
  scale_fill_manual(values = custom_colors, labels = bike_type_names) +  
  theme_minimal() +
  theme(
    legend.position = "right",  
    axis.text.x = element_text(angle = 45, hjust = 1)  
  )
```

### From Theft to Recovery

In an urban city like Toronto people are increasingly turning to
bicycles for convenience and sustainability, yet they face the threat of
losing their bikes. The significant disparity between stolen bikes and
recovered ones reflects the lack of infrastructure, resources, and
efficient tracking systems for bicycle theft. The 1% recovery rate of
bicycles becomes a rallying cry for activists and urban planners who
promote cycling as an eco-friendly alternative to cars. In response to
this challenge, the City of Toronto has introduced initiatives like
StreetSmartsTO, that is a program designed to support the city’s growing
cycling community. Although these efforts are a step in the right
direction, it is evident that there is still a long way to go in
creating a safer environment for cyclists. City also encourages
residents to register their bikes with the Toronto Police Service’s bike
registration program. While registering a bike can help track it in case
of theft, the reality is that this alone won’t solve the problem. In
order to see real change, Toronto will need to invest in smarter
tracking technologies and more secure bike parking options

This analysis suggests that Toronto Police and Biker's Association
should pay more attention to Downtown area in order to reduce the city’s
overall bike theft rate. The city should prioritize installation of
enough locking stations in well lit areas to provide adequate visibility
and deter criminals. As an individual, one must invest in a good solid
lock and register their bike with its serial number to save the trouble
if they loose their bike in future.